---
title       : rCharts
author      : Ramnath Vaidyanathan
framework   : minimal
highlighter : prettify
hitheme     : twitter-bootstrap
mode        : selfcontained
github      : {user: rcharts, repo: howitworks, branch: gh-pages}
widgets     : [disqus, ganalytics]
assets:
  css: 
    - "http://fonts.googleapis.com/css?family=PT+Sans"
    - "http://odyniec.net/articles/turning-lists-into-trees/css/tree.css"
---

# How rCharts Works?
#### Part 1

<!-- AddThis Smart Layers BEGIN -->
<!-- Go to http://www.addthis.com/get/smart-layers to customize -->
<script type="text/javascript" src="//s7.addthis.com/js/300/addthis_widget.js#pubid=ra-4fdfcfd4773d48d3"></script>
<script type="text/javascript">
  addthis.layers({
    'theme' : 'transparent',
    'share' : {
      'position' : 'left',
      'numPreferredServices' : 5
    }   
  });
</script>
<!-- AddThis Smart Layers END -->

```{r setup, echo = F, cache = F, message = F}
require(knitr)
opts_chunk$set(results = 'asis', comment = NA, tidy = F)
```

<a href="http://prose.io/#{{page.github.user}}/{{page.github.repo}}/edit/gh-pages/index.Rmd" class="button icon edit">Edit Page</a>

This is a multi-part series on how rCharts works. My objective is to show you how easy it is to integrate a javascript visualization library into rCharts, and take advantage of a single-unified interface and other functionalities. 

In this tutorial, I will go over the steps to integrate [uvCharts](http://imaginea.github.io/uvCharts/), a new-entrant to the rapidly expanding set of wrapper libraries to [d3js](http://d3js.org). Specifically, we will recreate this barchart, which was created using uvCharts.


<iframe
  style="width: 100%; height: 600px"
  src="http://jsfiddle.net/RR8Ub/1/embedded/result,resources,js,html">
</iframe>

Browse through the `resources`, `javascript` and `html` tabs in the jsFiddle to get a sense of how the chart works.  rCharts will need to create each of these sections, and we will go through them one by one.

### Resources

The resources required by a chart include the external javascript `<script src = "...js">` and css files `<link href="...css" rel="stylesheet">` that are included either in the head of the html document or at the end. rCharts uses a convention over configuration approach, and so the resources are specified using a `config.yml` file as shown below.

The [config](libraries/widgets/uvcharts/config.yml) file basically provides paths to the js/css files, both locally and  online from a cdn. The `jshead` key specifies that these js files need to be included in the head of the document.

```yaml
uvcharts:
  jshead: [js/d3.v3.min.js, js/uvcharts.js]
  cdn:
    jshead:
      - "http://cdnjs.cloudflare.com/ajax/libs/d3/3.2.2/d3.v3.min.js"
      - "http://imaginea.github.io/uvCharts/js/uvcharts.js"
```

At this point, the `uvCharts` library folder will look like this

<ul class="tree" id="tree">
  <li>js
    <ul>
      <li>d3.v3.min.js</li>
      <li class="last">uvcharts.js</li>
    </ul>
  </li>
  <li class="last">config.yml</li>
</ul>

<br/>

### Javascript

Our next step is to split the javascript for the chart into a [mustache](http://mustache.github.io/) layout, and then populate it using a `json` payload processed by R.  

<iframe
  style="width: 100%; height: 600px"
  src="http://jsfiddle.net/RR8Ub/1/embedded/js">
</iframe>

--- .RAW

By default, `rCharts` bundles all data and parameters into a single json variable `chartParams`. In addition to `chartParams`, it also sends `chartId` to the layout. So, we begin by replacing all data in the javascript with a placeholder `{{{ chartParams }}}`.  rCharts uses the `whisker` package to replace `{{{ chartParams }}}` with the json payload.

#### Layout

```html
<script>
  var graphdef = {{{ chartParams }}}
  var config = {
    meta: {
      position: "#{{ chartId }}"
    }
  }
  var chart = uv.chart(graphdef.type, graphdef, config)
</script>
```

Following rCharts conventions, we save this layout to `layouts/chart.html`.
At this point our [uvCharts](libraries/widgets/uvCharts) library folder is going to look like this 

<ul class="tree" id="tree">
  <li>js
    <ul>
      <li>d3.v3.min.js</li>
      <li class="last">uvcharts.js</li>
    </ul>
  </li>
  <li>layouts
    <ul>
      <li class="last">chart.html</li>
    </ul>
  <li class="last">config.yml</li>
</ul>

<br/>

---

#### Payload

Different d3js wrapper libraries tend to use different data structures. Our goal with `rCharts` is to provide a single-unified interface for specifying a chart, so that a user does not need to worry about the underlying format used by a specific library. This can be easily achieved by writing functions that transform data (usually a `data.frame`) to different formats, as required by the library.

Shown below is the dataset that we will be working with. We need to convert it into an R object, that when converted to JSON will look like the object `dataset` seen in the `javascript` shown above.


```{r results='markup'}
hair_eye_male <- subset(as.data.frame(HairEyeColor), Sex == "Male")
head(hair_eye_male)
```

We will now write a data transformation function in R that takes `x`, `y`, `data` and `group` as inputs, and returns a data structure that when converted to JSON will match what we need. The basic idea behind `make_dataset` is to (a) rename the `x`, `y` columns to `name`, `value`, (b) split the `data` based on `group` and (c) convert each piece into a JSON array.


```{r}
make_dataset <- function(x, y, data, group = NULL){
  require(plyr)
  dat <- rename(data, setNames(c('name', 'value'), c(x, y)))
  dat <- dat[c('name', 'value', group)]
  if (!is.null(group)){
    dlply(dat, group, toJSONArray, json = F)
  } else {
    list(main = toJSONArray(dat, json = F)) 
  }
}
```

You can run the code below to check that it produces the same `dataset` variable seen in the javascript above.

```{r eval = F}
dataset = make_dataset('Hair', 'Freq', hair_eye_male, group = 'Eye')
cat(RJSONIO::toJSON(dataset))
```

--- #mychart

### Chart Creation

It is now time to create the chart from R. We initialize the chart as an instance of `rCharts`, use the `setLib` method to point to the location of the library folder, and then pass the data payload using the `set` method. 

```{r}
library(rCharts)
u1 <- rCharts$new()
u1$setLib("libraries/widgets/uvcharts")
u1$set(
  type = 'Bar',
  categories = names(dataset),
  dataset = dataset,
  dom = 'chart1'
)
u1
```

For the above code to work, you need to make sure that you have [downloaded](https://github.com/rcharts/howitworks/zipball/gh-pages) this repo, and are referring to the correct path to the `uvCharts` library.

Alternately, you can point to the online version of the library `http://rcharts.github.io/howitworks/libraries/widgets/uvcharts` in `setLib`, and not have to download anything!!

You could save all this code in an file named [myChart.R](assets/code/mychart.R) and then publish the chart along with the code by running the code below. The published chart can be seen [here](http://rcharts.io/viewer/?3b61ed669d7f15156520)

```r
u2 <- create_chart('assets/code/myChart.R')
u2$publish('My Chart')
```

For many libraries rCharts already simplifies the steps.  Similarly, you can simplify the chart creation interface further, by wrapping everything into a `uPlot` function

```{r}
uPlot <- function(x, y, data, group = NULL, ...){
  dataset = make_dataset(x = x, y = y, data = data, group = group)
  u1 <- rCharts::rCharts$new()
  u1$setLib("libraries/widgets/uvcharts")
  u1$set(
    categories = names(dataset),
    dataset = dataset,
    ...
  )
  return(u1)
}
```

Let us create a `StackedBar` chart using `uPlot`

```{r stackedbar}
uPlot("Hair", "Freq", 
  data = hair_eye_male, 
  group = "Eye",
  type = 'StackedBar'
)
```

So you can see for yourself, how easy it is to integrate a javascript visualization library into `rCharts`. 

The [uvCharts](http://imaginea.github.io/uvCharts/documentation.html) API provides more functionality to customize the chart. While, we can use nested lists with the `set` method to access this functionality, it would be cleaner to write custom methods that follow the unified interface and make things easier. We will go over this in Part 2 of this series.

### Acknowledgements

I would like to thank [Kenton Russell](http://github.com/timelyportfolio) for his valuable suggestions and comments, which helped improve this post.


<div id='disqus_thread'></div>

