---
title       : rCharts
author      : Ramnath Vaidyanathan
framework   : minimal
highlighter : prettify
hitheme     : twitter-bootstrap
mode        : selfcontained
github      : {user: rcharts, repo: howitworks, branch: gh-pages}
widgets     : [disqus, ganalytics]
assets:
  css: 
    - "http://fonts.googleapis.com/css?family=PT+Sans"
    - "../assets/css/app.css"
    - "../assets/css/gh-buttons.css"
url: {lib: ../libraries}
---

## Micropolar with rCharts

[Micropolar](http://micropolar.org) is a polar chart library made with `d3.js`. I saw the announcement on twitter by [@d3visualization](https://twitter.com/d3visualization) and could not resist the urge to integrate it with rCharts.

I should say that `micropolar` is extremely well designed for reusability.  The author Chris Viau is also a coauthor of the very good book [Developing a D3.js Edge](http://bleedingedgepress.com/our-books/developing-a-d3-js-edge/) that details how to make `d3.js` charts reusable.  Well designed reusable charts integrate easily with `rCharts`.

### Getting Started

We need a couple of helper functions to transform a data frame to the format required by micropolar, and to create a polar plot.

```{r message = F, echo = F, cache = F}
library(knitr)
opts_chunk$set(results = 'asis', comment = NA, tidy = F)
```

```{r}
# Convert data frame to format required by micropolar
make_dataset = function(x, y, data = data){
  require(rCharts)
  toJSONArray2(data[c(x, y)], json = F, names = F)
}

# Create a polar plot given data and type
polarPlot <- function(data, type, ...){
 u <- rCharts$new()
 u$setLib('http://rcharts.github.io/howitworks/libraries/widgets/micropolar')
 if (!missing(data)) u$set(data = data)
 u$set(type = type, ...)
 return(u)
}
```

Let us define some data that we can use

```{r}
dat = data.frame(
  x = c(60, 180, 270, 360),
  y = c(5, 2, 3, 4)
)
```


### DotPlot

```{r dotPlot, message = F}
require(rCharts)
options(RCHART_HEIGHT = 400, RCHART_WIDTH = 400)
polarPlot(
  data = make_dataset('x', 'y', dat),
  type = 'dotPlot'
)
```

### LinePlot

We are working on some useful examples combining R data with the other `micropolar` chart types.  For now, let's look at each chart type using the defaults specified by `micropolar`.

```{r linePlot}
polarPlot(type = 'linePlot')
```

### AreaChart

```{r areaChart}
polarPlot(type = 'areaChart')
```

### BarChart

```{r barChart}
polarPlot(type = 'barChart')
```

### Clock

```{r clock}
polarPlot(type = 'clock')
```

### Acknowledgements

I would like to thank [TimelyPortfolio](http://github.com/timelyportfolio) for adding most of the examples in this post!

<div id='disqus_thread'></div>

<style>
iframe {
  height: 410px;
}
</style>





